---
title: "FinalReport"
author: "Steven Garrity + Richard Palmer"
date: "10/26/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r child = 'dataload.Rmd'}
```

# Introduction

The purpose of this project is to study a dataset of 
`r nrow(beers_raw)` beers produced by `r nrow(breweries_raw)`
breweries in the United States.  We have data on the style, ABV (alcohol by volume)
and IBU (international bitterness units) of each beer, and we have the city
and state for each brewery.  The goal is to answer some specific questions 
about the data and to look for relationships or inferences which may be of
interest to a hypothetical sponsor, Budweiser.

```{r child = 'question1.Rmd'}
```

```{r child = 'question2.Rmd'}
```

```{r child = 'question3.Rmd'}
```

```{r child = 'question4.Rmd'}
```

```{r child = 'question5.Rmd'}
```

```{r child = 'question6.Rmd'}
```

```{r child = 'question7.Rmd'}
```

```{r child = 'question8.Rmd'}
```

#```{r child = 'question9.Rmd'}
#```

