---
title: "FinalReport"
author: "Steven Garrity and Richard Palmer"
date: "26 October 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r child = 'dataload.Rmd'}
```

# Prepared for

Prepared for CEO and CFO of Budweiser, a subsidiary of Anheuser-Busch.  

Case Study 1 project for DS6306 Section 408  Fall 2019  Dr. Bivin Sadler

# Introduction

The purpose of this project is to study a dataset of 
`r nrow(beers_raw)` beers produced by `r nrow(breweries_raw)`
breweries in the United States.  We have data on the style, ABV (alcohol by volume)
and IBU (international bitterness units) of each beer, and we have the city
and state for each brewery.  The goal is to answer some specific questions 
about the data and to look for relationships or inferences which may be of
interest to a hypothetical sponsor, Budweiser.

```{r child = 'question1.Rmd'}
```

```{r child = 'question2.Rmd'}
```

```{r child = 'question3.Rmd'}
```

```{r child = 'question4.Rmd'}
```

```{r child = 'question5.Rmd'}
```

```{r child = 'question6.Rmd'}
```

```{r child = 'question7.Rmd'}
```

```{r child = 'question8.Rmd'}
```

```{r child = 'question9.Rmd'}
```

# Conclusion

It is clear that the craft beer segment of the industry has taken hold across the country.
There are moderate but clear differences between states in the characteristics of beers.

Examining available measurements of craft beers, we have found statistically significant
differences in ABV and IBU between different classes of ales.

Looking at the number of breweries in each state, along with population, we have found
eleven states that represent a potentially underserved market.  Considering the beers 
that are currently marketed in those states, we have identified four distinct groups of
states with similar tastes.  Based on the characteristics of beers that are popular in 
those states, we believe three new labels with appropriate styles could do well in the
underserved market.  Here is our recommendation:

```{r}
library(knitr)
library(kableExtra)
market <- c('GA,FL,NY','MO,TX,MD,IL,TN,OH,CA','NJ')
style <- c('IPA','Higher Alcohol Ale','Lower Alcohol Ale')
recommendation <- data.frame(market,style)
colnames(recommendation) <- c('Target Marget','Suggested Beer Style')
kable(recommendation,caption='Stratety for Underserved Market') %>%
    kable_styling(bootstrap_options=c('striped'))
```

