---
title: "FinalReport"
author: "Steven Garrity + Richard Palmer"
date: "10/6/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r child = 'dataload.Rmd'}
```

# Introduction

```{r}
beer_count <- nrow(beers_raw)
brewery_count <- nrow(breweries_raw)
style_count <- nlevels(beers_raw$Style)
```

We were presented with data on `r beer_count` beers produced by `r brewery_count` breweries in the United States.

For each brewery, we have its Name, and the city and state it is located in.

For each beer, we have its style as well as numeric values for ABV (alcohol by volume),
IBU (international bitterness units), and the serving size in ounces.

In this dataset, there were `r style_count` different styles of beer represented.

There were beers produced in every state of the nation, as well as in the District of Columbia.


```{r child = 'question1.Rmd'}
```

```{r child = 'question2.Rmd'}
```

```{r child = 'question3.Rmd'}
```

```{r child = 'question4.Rmd'}
```

```{r child = 'question5.Rmd'}
```

```{r child = 'question6.Rmd'}
```

```{r child = 'question7.Rmd'}
```

```{r child = 'question8.Rmd'}
```

```{r child = 'question9.Rmd'}
```

