---
title: "question3"
author: "Steven Garrity"
date: "10/9/2019"
output: html_document
summary: "We found that IBU and ABV were the only features with missing values. ABV had XX NAs and IBU had 1005 NAs. All of the ABV NAs also had NA IBU. We chose to fill ABV values with the median ABV from beers of the same type. To fill missing values in IBU, we fit a regression tree to observations without missing values and then used that model to predict each missing IBU value."
---

```{r question3, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question 3: Address the missing values in each column.

Data Exploration:
```{r}
########################################################################
# Data Exploration
########################################################################
n_cols = dim(merged)[2]
for (i in 1:n_cols)
{
  print(c(colnames(merged[i]), sum(is.na(merged[i]))))
}
# It appears that ABV and IBU are the only columns with missing values.

# Extract the row indices of the missing values
missingIBUindex <- which(is.na(merged$IBU))
missingABVindex <- which(is.na(merged$ABV))
allNAindex <- c(missingIBUindex,missingABVindex)
length(missingIBUindex)
length(unique(allNAindex)) 
```

Fill ABV missing values:
```{r}
########################################################################
# Fill ABV missing values
########################################################################

# Find the style of each beer with a missing value, calculate the median ABV for that style of beer, and fill the missing value with the median for that style. Filled IBU values are placed into the "merged" dataset as merged$IBUfill.
merged$ABVfill <- merged$ABV
for (i in 1:length(missingABVindex))
{
  merged$ABVfill[missingABVindex[i]] <- median(merged$ABV[merged$Style == merged$Style[missingABVindex[i]]],na.rm = TRUE)
}

# sanity check to make sure that we filled all of the missing values:
which(is.na(merged$ABV))
which(is.na(merged$ABVfill)) # looks like all values were filled...we are good to proceed
```

Fill IBU missing values: ***note that I use "set.seed(8)" to ensure that the model predictions are always the same. My understand is that this consistency should ensure that any downstream dependency shouldn't break. Filled IBU values are placed in the merged data frame as merged$IBUfill.
```{r}
########################################################################
# Fill IBU missing values
########################################################################
# Let's try a regression tree (CART)...throw the kitchen sink at it...:)
# library(rpart)
# library(rpart.plot)
# ?rpart

# Get data into the right format:
forTraining_IBU_rtree <- merged[-missingIBUindex,] %>% 
  select(ABV,Style,Brewery_name,City,State,Ounces,IBU)

forFilling_IBU_rtree <- merged[missingIBUindex,] %>% 
  select(ABV,Style,Brewery_name,City,State,Ounces,IBU)

# set up train/test split
set.seed(8)
splitPerc = .75
trainIndices = sample(1:dim(forTraining_IBU_rtree)[1],round(splitPerc * dim(forTraining_IBU_rtree)[1]))
train = forTraining_IBU_rtree[trainIndices,]
test = forTraining_IBU_rtree[-trainIndices,]

# train the model
rtree <- rpart(IBU ~ ABV+Style, 
               data = train,
               method = "anova",
               control = list(cp = 0.0001, xval = 1000))

plotcp(rtree) # visualize relative error by tree size
# prp(rtree) # visualize the tree structure

######
# evaluate model performance
rtree.pred = predict(rtree, test)
rtree.sse = sum((rtree.pred - test$IBU)^2)
rtree.mse = mean((rtree.pred - test$IBU)^2)
rtree.rmse = sqrt(mean((rtree.pred - test$IBU)^2))
rtree.sse
rtree.mse
rtree.rmse

test$predictions <- rtree.pred

test %>% ggplot(aes(x=IBU, y=predictions)) +
  geom_point(col='navyblue') +
  geom_smooth(method='lm', col="darkgrey") +
  labs(title='Regression Tree Model Performance', x="Actual IBU", y="Predicted IBU")
######

### sanity check before updating the "merged" data frame
# merged$IBU[missingIBUindex[1]]
# merged$IBU[missingIBUindex[2]]
# predict(rtree, forFilling_IBU_rtree[1,])
# predict(rtree, forFilling_IBU_rtree[2,])

merged$IBUfill <- merged$IBU
merged$IBUfill[missingIBUindex] <- predict(rtree, forFilling_IBU_rtree)

### complete the sanity check:
# merged$IBU[missingIBUindex[1]]
# merged$IBU[missingIBUindex[2]]
```

