---
title: "question6"
output: html_document
---

```{r question6, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Question 6
Comment on the summary statistics and distribution of the ABV variable.
```{r}
summary(merged$ABVfill)
sd(merged$ABVfill)
summary(merged$ABV)
sd(merged$ABV,na.rm=TRUE)

```

The mean is slightly higher than the median, which would indicate a right skewed distribution,
one with a long tail of high values.

```{r}
merged %>%
  ggplot(aes(x=ABVfill)) + geom_histogram(aes(y=..density..),bins=27) + 
  ggtitle('Histogram of ABV values (missing values infilled)') +
  geom_vline(aes(xintercept=median(ABVfill)),color='blue',linetype='dashed',size=1) +
  geom_text(mapping=aes(x=median(ABVfill),y=40,label="median"),color="blue",angle=90,vjust=-0.4,show.legend=FALSE) +
  geom_vline(aes(xintercept=mean(ABVfill)), color='red',linetype='dotted',size=1) +
  geom_text(mapping=aes(x=mean(ABVfill),y=40,label="mean"),color="red",angle=90,vjust=-0.4,show.legend=FALSE) +
  geom_density(alpha=.2,fill="#FF6666")
```

The histogram bears this out, as does a box plot

```{r}
merged %>%
  ggplot(aes(x="",y=ABVfill)) +
  geom_boxplot() +
  xlab("") + ylab("") +
  ggtitle('ABV values') +
  #geom_hline(yintercept=0.07,color="blue",show.legend=FALSE) +
  #geom_text(mapping=aes(y=0.07,label="mean"),color="blue",vjust=-1,show.legend = FALSE) +
  theme(aspect.ratio = 2.0) +
  coord_cartesian(ylim=c(0,0.15))
```

To investigate whether the ABV values are normally distributed, we do
a Q-Q plot.  This is a plot of the quantiles of the actual data against
what would be expected from a normal distribution.  If the data do
follow a normal distribution, the plot would show a straight line.  The
more the plot deviates from a straight line, the less closely 
the data follow a normal distribution.

```{r}
merged %>% 
  ggplot(aes(sample=ABVfill)) +
  stat_qq() + stat_qq_line() + 
  ggtitle('Quantile-Quantile plot of (infilled) ABV data')
```

From the upward curve in this plot, we infer that the data are somewhat skewed.
This confirms what the histogram shows.  The apparent discontinuities at the
ends of the curve are an indication that the ABV values have a few more extreme
values than would be expected from a normal distribution.  This confirms what
was shown in the box plot.
