---
title: "question4"
output: html_document
---

```{r question4, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

library(usmap)
library(ggthemes)

```

# Question 4
Compute the Median alcohol content for all the beers in each state.

```{r}
# drop any observations without valid ABV values, group by state, and compute the median
abv_df <- merged %>% group_by(State) %>% 
  summarize(median_abv = median(ABV,na.rm=TRUE),median_abvfill = median(ABVfill)) 

# order by median ABV
abv_df$State <- factor(abv_df$State,level=abv_df$State[order(abv_df$median_abvfill,decreasing=TRUE)])
```

Here are the original data, with missing observations dropped:

```{r}
# ..and plot
abv_df %>% ggplot(aes(x=State,y=median_abv)) + 
  geom_col(color="#13294b",fill="#C8102E") + 
  theme(aspect.ratio = 0.2,axis.text=element_text(size=4)) +
  ylab('Median ABV') + 
  ggtitle('Median ABV by state (missing values dropped)')

```

and here is the same plot, with missing values infilled.
Note that the ordering of states is the same for these two graphs.

```{r}
abv_df %>% ggplot(aes(x=State,y=median_abvfill)) + 
  geom_col(color="#13294b",fill="#C8102E") + 
  theme(aspect.ratio = 0.2,axis.text=element_text(size=4)) +
  ylab('Median ABV') + 
  ggtitle('Median ABV by state (missing values infilled)')
```

Here is the same data, with each state colored by the its median ABV value.

```{r}
# change up a bit to make a map
abv_df$State <- str_trim(abv_df$State, "left")

colnames(abv_df)[1] = "abbr"
abv_df$abbr <- as.character(abv_df$abbr)

abv_by_state_tomap <- merge(abv_df, states_and_dc, by="abbr")

plot_usmap(regions = c("states"), data = abv_by_state_tomap,
  values = "median_abvfill", theme = theme_map(), labels = FALSE,
  label_color = "black") +
  scale_fill_continuous(name = "ABV", low="ivory", high="#c8102e") +
  theme(legend.position="right") +
  labs(title = 'Median ABV of beers by state, infilled data')

```

Compute the median bitterness unit for all the beers in each state
```{r}
# drop any obsvervations without valid IBU values, roup by state, and compute the median
ibu_df <- merged %>% group_by(State) %>% 
  summarize(median_ibu = median(IBU,na.rm=TRUE),median_ibufill=median(IBUfill)) 

# order by median IBU
ibu_df$State <- factor(ibu_df$State,level=ibu_df$State[order(ibu_df$median_ibufill,decreasing=TRUE)])
```

Here are the original data, with missing observations dropped:

```{r}
# ..and plot
ibu_df %>% ggplot(aes(x=State,y=median_ibu)) +
  geom_col(color="#13294b",fill="#C8102E") + 
  theme(aspect.ratio = 0.2,axis.text=element_text(size=4)) +
  ylab('Median IBU') + 
  ggtitle('Median IBU by state (missing values dropped)')
```

and here is the same plot, with missing values infilled.
Note that the ordering of states is the same for these two graphs.

```{r}
ibu_df %>% ggplot(aes(x=State,y=median_ibufill)) +
  geom_col(color="#13294b",fill="#C8102E") + 
  theme(aspect.ratio = 0.2,axis.text=element_text(size=4)) +
  ylab('Median IBU') + 
  ggtitle('Median IBU by state (missing values infilled)')
```

Here is the same data, with each state colored by the its median IBU value.

```{r}
# change a bit to make a map
ibu_df$State <- str_trim(ibu_df$State, "left")

colnames(ibu_df)[1] = "abbr"
ibu_df$abbr <- as.character(ibu_df$abbr)

ibu_by_state_tomap <- merge(ibu_df, states_and_dc, by="abbr")

plot_usmap(regions = c("states"), data = ibu_by_state_tomap,
  values = "median_ibufill", theme = theme_map(), labels = FALSE,
  label_color = "black") +
  scale_fill_continuous(name = "ABV", low="ivory", high="#13294b") +
  theme(legend.position="right") +
  labs(title = 'Median ABV of beers by state')
```























